Bounds, Semantics/Pragmatics-Divide, and Evolution

### 2016-08-01 paper/paper.tex
paper introduction modifications

\begin{document}
\maketitle
\tableofcontents
\section{Introduction}\label{sec:introduction}

\hl{some general sentence here}
In linguistic theorizing, it is common to draw a distinction between semantics and pragmatics. Broadly speaking, semantics concerns the truth-conditional content of expressions, whereas pragmatics deals with information beyond such literal meanings. For instance, the integration of context and the derivation of defeasible inferences. Thus, the information conveyed is seldom if ever solely determined by semantics, but rather by its pragmatic enrichment. The present investigation addresses the selection of particular semantics over others in light of pragmatics. %  the semantics-pragmatics distinction raises the challenge of justifying the former's structure.

Consider, for instance, the utterance {\em Bill read five books}. Truth-conditionally, this utterance does not set an upper-bound for the amount of books that Bill read; he may have read six, seven, or more books. Pragmatically, however, an addressee who assumes the speaker to be cooperative and knowledgeable may draw the (defeasible) inference that {\em Bill read no more than five books}. After all, had he read more, for example, six books, and the speaker had known this, she would have said {\em Bill read six books} instead. In this way, what is conveyed (pragmatically) goes beyond what is (literally) said \citep{grice:1975}.

Much research in semantics and pragmatics has been aimed at characterizing the information conveyed by classes of expressions, such as numerals, in terms of either domain, or their interplay. Thus, while there is disagreement on where their boundary lies, their distinction remains a core (if tacit) assumption in linguistics. An issue that has received little attention, however, is the justification of semantic structure in view of pragmatics. Prima facie, under a functionalist view, is is puzzling that {\em Bill read five books} does not semantically rule out its more informative or ``stronger'' alternatives. Wouldn't it serve speakers of a language better if weak(er) expressions such as {\em warm}, {\em or}, {\em some} or {\em big} were not truth-conditionally compatible with stronger alternatives such as {\em hot}, {\em and}, {\em all} and {\em huge}, respectively? After all, this is what an interlocutor's pragmatically enriched interpretation yields in the case of so-called scalar inferences. Once a distinction between semantics and pragmatics is drawn, similar questions can be posed for other types of pragmatic inferences, such as ignorance inferences as (\ref{ex:ignorance}), or manner inferences such as (\ref{ex:manner-1}) and (\ref{ex:manner-2}).

\begin{exe}
\ex\label{ex:ignorance} \begin{xlist}
	\ex Bill read at least five books
	\ex $\leadsto$ Bill read five or more books (but the speaker doesn't know how many)
\end{xlist}
\ex \label{ex:manner-1} \begin{xlist}
	\ex Mercader killed Trotsky.
	\ex $\leadsto$ Mercader killed Trotsky in a stereotypical way
    \end{xlist}
\ex \label{ex:manner-2} \begin{xlist} \ex Mercader caused Trotsky to die
		  \ex $\leadsto$ Mercader killed Trotsky in a non-stereotypical way
    \end{xlist}
\end{exe}

>>>>

\begin{document}
\maketitle
\tableofcontents
\section{The semantics-pragmatics divide}\label{sec:introduction}

In linguistic theorizing, it is common to draw a distinction between semantics and pragmatics. Broadly speaking, the former concerns the truth-conditional content of expressions, whereas the latter concerns information beyond such literal meanings. For instance, the integration of context and the derivation of defeasible inferences. Thus, the information conveyed by an utterance is seldom, if ever, solely determined by semantics, but rather by its pragmatic enrichment. For instance, truth-conditionally, {\em Bill read five books} does not set an upper-bound to the amount of books that Bill read; he may have read six, seven, or more books. However, the (defeasible) inference that {\em Bill read no more than five books} may be drawn on pragmatic grounds. After all, if Bill had read, for example, six books, and had the speaker known this, she would have said so. In this manner, what is conveyed (pragmatically) can go beyond what is (literally) said \citep{grice:1975}.

Much research in semantics and pragmatics is aimed at characterizing the information conveyed by classes of expressions such as numerals in terms of either domain, or their interplay. Thus, while there is disagreement on where their boundary lies, the distinction between semantics and pragmatics plays an important (if tacit) role in linguistics. However, an issue that has received little attention is the justification of semantic structure in light of pragmatics. Put differently, this issue concerns the selection and pervasiveness of certain semantic structures over others under consideration of the informational enrichment provided by pragmatics.

Prima facie, it is puzzling that {\em Bill read five books} does not semantically rule out its more informative or ``stronger'' alternatives. More pointedly, would it not serve language users better if weak(er) expressions such as {\em warm}, {\em or}, {\em some} or {\em big} were not truth-conditionally compatible with stronger alternatives such as, respectively, {\em hot}, {\em and}, {\em all} and {\em huge}? After all, analogously to the example sketched above, this is what a pragmatically enriched interpretation yields. Once a distinction between semantics and pragmatics is drawn, similar questions can be posed for other types of pragmatic inferences, such as ignorance inferences as (\ref{ex:ignorance}), or manner inferences such as (\ref{ex:manner-1}) and (\ref{ex:manner-2}).
